% SPDX-License-Identifier: CC-BY-NC-SA-4.0
%
% thank-you.tex -- thank you!
% Copyright (C) 2023  Jacob Koziej <jacobkoziej@gmail.com>

\pdfbookmark{Thank You}{thank-you}

\thispagestyle{empty}

\begingroup
\hspace{0pt}
\vfill
\noindent
This book wouldn't be possible if it were not for some very special
people.

\vspace{0.5em}

\noindent
Thank you to Rachel Kaufman for letting me find myself and not only
discover my passion for programming but for teaching as well.  The joy
your teaching has brought into my life is immeasurable, and I truly hope
to bring the same joy you've brought me to anyone I teach.

\vspace{0.5em}

\noindent
Ross Grosshart for the countless hours spent chatting on Zoom during the
COVID-19 pandemic about taking life one step at a time.  Talking with
you helped me get through the pandemic and figure out where I want to
take my career.

\vspace{0.5em}

\noindent
Dan Mezhiborsky for the mentorship and patience.  Your pushing has made
me a much better engineer and let me see that there's so much more to
life than studying at every waking moment.

\vspace{0.5em}

\noindent
Michael Giglia for having enough faith in me to allow me to break any
code I touch before improving it.  You've helped me hone in on what's
important while building bigger and better things.

\vspace{0.5em}

\noindent
Tara Menon, for helping me see the world in a positive light and for all
the hours we've spent working on improving my writing and discussing
pedagogy.  I'll always be grateful for the second chance you gave me
from my abysmal first impression and for encouraging me to speak up and
take on this book.

\vspace{0.5em}

\noindent
And with that, I present you: \emph{A Zoomer's Guide to C Programming}
\vfill
\hspace{0pt}
\endgroup

\cleardoublepage
